Spin canonical rings of log stacky curves

### 2015-07-08 curves-write-up/spin_cring.tex
field k -> \Bk, indices k -> \ell

\newcommand\nc{\newcommand}
\nc\on{\operatorname}
\nc\renc{\renewcommand}
\newcommand\ssec{\subsection}
\newcommand\sssec{\subsubsection}
\newcommand\BH{{\mathbb H}}
\newcommand\BN{{\mathbb N}}
\newcommand\BC{{\mathbb C}}
\newcommand\BF{{\mathbb F}}
\newcommand\BR{{\mathbb R}}
\newcommand\BQ{{\mathbb Q}}
\newcommand\BP{{\mathbb P}}
\newcommand\BZ{{\mathbb Z}}
\newcommand\BA{{\mathbb A}}

@@

Let $X$ be a smooth proper geometrically-integral algebraic curve of genus $g$ over a field $k$.
It is well known that the canonical sheaf $\Omega _X,$ with
associated canonical divisor $K_X$, determines the {\bf canonical
map } $\pi: X \rightarrow \BP_k^{g - 1}$. Then, the {\bf canonical
ring} is defined to be
\begin{align*}
	R(X, K_X) = \bigoplus_{d \geq 0} H^0(X, dK_X),
\end{align*}

@@

\begin{thm}
\label{thm:main}
Let $(\sx, \Delta, \halfcan)$ be a log spin curve over a perfect
field $k$, so that $\sx$ has signature $\sigma = (g; e_1, \ldots,
e_r; \delta)$.

Then the log spin canonical ring is generated as a $k$-algebra by 
elements of degree at most $e = \max(5, e_1, \ldots, e_r)$ with
relations generated in degrees at most $2e$,
so long as $\sigma$ does not lie in a finite list of exceptional
cases, as given in Table ~\ref{table:g-1-exceptional} for
signatures with $g = 1$ and Table ~\ref{table:g-0-exceptional} for
signatures with $g = 0$.
\end{thm}

\begin{rem}
\label{rem:main-weaker-assumptions}
In fact, the proof of Theorem ~\ref{thm:main} holds with $\Delta$ replaced by an arbitrary effective divisor of the coarse space.
 Furthermore, one may relax the assumption that $k$ is perfect. Instead, one only need assume that the stacky curve
is separably rooted, as described further in Remark ~\ref{rem:stack-formalism}.
\end{rem}

@@

For the remainder of this paper, fix an algebraically closed field $k$.
This is no restriction on generality, as generator and relation degrees are preserved under base change to the algebraic closure.

@@

\begin{defn}
\label{defn:stacky-curve}
A \textbf{stacky curve} $\sx$ over an algebraically closed field $k$ is the datum of a smooth proper connected scheme $X$ of dimension $1$, together with a finite number of closed points of $X$, $P_1, \ldots, P_r$, called {\bf stacky points}, with {\bf stabilizer orders} $e_1, \ldots, e_r \in \BZ_{\geq 2}.$ The scheme $X$ associated to a stacky curve $\sx$ is called the {\bf coarse space} of $\sx$.

@@

The results of this paper can be easily couched in terms of the language of stacks. If one works over an arbitrary field $k$ (which need not be algebraically closed) one can extend Theorem ~\ref{thm:main} to hold in the case that the stacky curve $\sx$ is tame and separably rooted, i.e.,~ the residue field of each
of the stacky points is separable.

@@

\begin{defn}
\label{defn:div-ex}
Let $\sx$ be a stacky curve over $k$ with coarse space $X$ of genus $g$ and stacky points $P_1, \ldots,
P_r$ with stabilizer orders $e_1, \ldots, e_r \in \BZ_{\geq 2}$.
Then, we notate
\[
	\di \sx = \left(\bigoplus_{P\notin \{P_1, \ldots, P_r\}} \langle 
	P \rangle \right) \oplus \left(\bigoplus_{i = 1}^r \left \langle 
	\frac{1}{e_i}P_i \right \rangle \right) \subseteq \BQ \otimes \di X.
\]
\end{defn}

@@

\begin{defn}
\label{defn:h0-stacky}
Recall the standard notation 
$$\sco(D) = \{f \in k(X)^\times : \di f + D \geq 0 \}.$$
Let $\sx$ be a stacky curve with coarse space $X$.
If $D \in \di \sx$ is a Weil divisor, then we define
\begin{align*}
	H^0(\sx, D) &:= H^0(X, \lfloor D \rfloor)\\
	H^0(\sx, \sco(D)) &:= H^0(\sx, D)\\
	h^0(\sx, \sco(D)) &:= \dim_k H^0(\sx, \sco(D))
\end{align*}
If $R$ is a graded ring, then we let $(R)_d$ refer to the $d^\text{th}$
graded component of $R$.

@@

\begin{defn}
\label{defn:monomial-degree}
Let $k[x_1$, $\ldots$, $x_n]$ be a graded polynomial
ring with $\deg x_i = d_i$ and let $\alpha := \prod_{i = 1}^{n} x_{i}^{f_i} \in
k[x_{1}$, $\ldots$, $x_{n}]$ be a monomial. Then
we define the {\bf degree} of $\alpha$ to be
\[
	\deg \alpha := \sum_{i = 1}^{n} d_i f_i.
\]
\end{defn}

\begin{defn}
\label{defn:grevlex}
The {\bf graded reverse lexicographic order}, or {\bf grevlex} $\prec_{\text{grevlex}}$ is defined as follows.
If $\alpha := \prod_{i = 1}^{n} x_{i}
^{f_i}$ and $\beta := \prod_{i = 1}^{n} x_{i}^{f_i'}$ are
monomials in $k[x_{1}$, $\ldots$, $x_{n}]$, then $\alpha
\succ_{\text{grevlex}} \beta$ if either

@@

\begin{defn}
\label{defn:block-order}
The \textbf{block term order} is defined as follows. Let $k[y_{1}, \ldots, y_{m}]$ and $k[x_{1}, \ldots,
x_{n}]$ be weighted polynomial rings with $\deg y_i = c_i, \: \deg x_i = d_i.$ Further assume we are given existing term
orders $\prec_y$ and $\prec_x$. Let $\alpha :=
\prod_{j = 1}^{m} y_{j}^{h_i}$ $\prod_{i = 1}^{n} x_{i}
^{f_i}$ and $\beta := \prod_{j = 1}^{m} y_{j}^{h_i'}
\prod_{i = 1}^{n} x_{i}^{f_i'}$ be monomials in $k[y_{1},
\ldots, y_{m}] \otimes$ $k[x_{1}, \ldots, x_{n}]$.
Let $\alpha_y := \prod_{j = 1}^{m} y_{j}^{h_i}$ be the
part of $\alpha$ in $k[y_{1}, \ldots, y_{m}]$ and
likewise with $\alpha_x$, $\beta_y$, and $\beta_x$.

In the \textbf{(graded) block} (or \textbf{elimination}) term
ordering on $k[y_{d_1', 1}, \ldots, y_{d_m', m}] \otimes k[x_{d_1, 1},
\ldots, x_{d_n, n}]$, we define $\alpha \succ \beta$ if
  \begin{enumerate}
	\item[(i)] $\deg \alpha  > \deg \beta$ or
  \item[(ii)] $\deg \alpha = \deg \beta$ and $\alpha_y \succ_y \beta_y$ or
	\item[(iii)] $\deg \alpha = \deg \beta$ and $\alpha_y = \beta_y$ and $\alpha_x \succ_x \beta_x$.
  \end{enumerate}
\end{defn}

@@

\begin{defn}
\label{defn:initial-term}
Let $\prec$ be an ordering on $k[x_{1}$, $\ldots$, $x_{n}]$, 
with $\deg x_i = d_i$,
and let $f \in k[x_{1}, \ldots, x_{n}]$ be a homogeneous
polynomial. The \textbf{initial term} $\initial(f)$ of $f$
is the largest monomial in the support of $f$ with respect to
the ordering $\prec$. Furthermore, we set $\initial(0) := 0$.
\end{defn}

\begin{defn}
\label{defn:initial-ideal}
Let $I$ be a homogeneous ideal of $k[x_{1}$, $\ldots$,
$x_{n}]$. Then the {\bf initial ideal} $\initial(I)$ of $I$ is
the ideal generated by the initial terms of homogeneous polynomials
in $I$:
\[
	\initial(I) := \langle \initial(f) \rangle_{f \in I}
\]
\end{defn}

@@

\begin{defn}
\label{defn:grobner-basis}
Let $I$ be a homogeneous ideal of $k[x_{1}$, $\ldots$,
$x_{n}]$. A \textbf{Gr\"obner basis} for $I$, also known as a
\textbf{standard basis} for $I$, is a set of elements in $I$
such that their initial terms generate the initial
ideal of $I$.
\end{defn}

@@

Let $R'$ be the canonical ring associated to $\sx'$. Define $(\sx', \halfcan', J)$
to be {\bf admissible} if $R'$ admits a presentation
\begin{align*}
	R' \cong \left( k[x_1, \ldots, x_m] \otimes k[y_{i, e_i'}]_{i \in J} \right)/I'
\end{align*}

@@

\begin{example}
\label{eg:base-1-0}
Let $(\sx',0, \halfcan')$ be a log stacky curve of genus $g = 1,$ with $\halfcan' = 0$.
Counting dimensions, we see $h^0(\sx, dL')=1$ for all $d\in \mathbb{Z}_{\ge 0}$ so it is immediately clear that $R_{\halfcan'} \cong k[x]$ with $x$ a generator in degree 1.
\end{example}

@@

Note that $R_{L_X} \cong k[x_1, x_2]$ where $\deg x_1 = \deg x_2 = 1.
$ To see this, observe that we will need two generators in degree $1$
because $h^0(X, L_X) = 2$ by Riemann--Roch. Because $L_X$ is very 
ample, we have that $R_{L_X}$ is generated in degree 1. To
conclude, note that $R_{L_X}$ does not have any relations. If there
exists some relation, then $\dim \proj R_{L_X} < 1$. This would
contradict the fact that $L_X$ is very ample. Thus, $\proj R_{L_X}
\cong X$ which has dimension 1.

Next, we construct generators and relations for $R_L$ using those
of $R_{L_X}$. Note that we have a natural inclusion $\iota: R_{L_X}
\hookrightarrow R_L$ induced by the inclusions $H^0(\sx, dL_X)
\hookrightarrow H^0(\sx, dL)$ for each $d \geq 0$. By Riemann Roch,
we see there is some element $y_{1, 3} \in (R_L)_3$ with $\ord_{P_1}
(y_{1, 3}) = -1$, not in the image of the inclusion $\iota$. We
claim that there exist $a_1, a_2 \in k$ and a degree $4$ polynomial
$f(x_1, x_2) \in k[x_1, x_2]$ such that
\begin{align*}
	R_L \cong k[x_1, x_2, y_{1, 3}] / (a_1 x_1 y_{1, 3} + a_2 x_2 y_{1, 3}
	+ f(x_1,x_2))
\end{align*}

@@

\noindent
are surjective for $d \geq 4$. A relation of the form $a_1 x_1 y_{1,
3} + a_2 x_2 y_{1, 3} + f(x_1, x_2) = 0$ must exist because $h^0(\sx,
4L) - h^0(\sx, 4L_X) = 1,$ but $x_1 y_{1, 3}$ and $x_2 y_{1, 3}$
define two linearly independent elements with nontrivial image in
the 1-dimensional vector space $H^0(\sx, 4L) / H^0(\sx, 4L_X).$ So,
we obtain a surjection
\begin{align}
\label{align:surjection}
	k[x_1, x_2, y_{1, 3}] / (a_1 x_1 y_{1, 3} + a_2 x_2 y_{1, 3} + f(x_1, x_2)) 
	\rightarrow R_L.
\end{align}

To complete the example, it suffices to show there are no
additional relations. First, note that $a_1 x_1 y_{1, 3} + a_2 x_2
y_{1, 3} + f(x_1, x_2)$ is irreducible because there are no
relations among $x_1, x_2$ and $y_{1, 3}$ in lower degrees. Hence,
$k[x_1, x_2, y_{1, 3}] / (a_1 x_1 y_{1, 3} + a_2 x_2 y_{1, 3} +
f(x_1, x_2))$ is integral and is 2-dimensional. Thus, the map
\eqref{align:surjection} defines a surjection from an integral
2-dimensional ring to a 2-dimensional ring. Therefore, it is an 
isomorphism.
\end{example}

@@

For $d = 0, 1, 2, \ldots$ we have
\[
	\dim_k H^0 (\sx', d K_{\sx'}) = 1, 0, 0, 1, 0, 1, 1, 2, 1, 1, 2, 1, 3, 2, 3, \ldots
\]

@@

Let $I$ be the ideal generated by these relations in
$k[x_{7, 2}, x_{7, 1}, x_{5, 1}, x_{3, 1}]$. Under grevlex with
$x_{3,1} \prec x_{5,1} \prec x_{7,1} \prec x_{7,2}$, the initial
ideal of $I$ is
\[
	\initial(I) = \langle x_{7, 2}^2, x_{5, 1}^2 \rangle
\]

@@

To demonstrate that these two relations generate all 
relations, we show that $\dim_k (R')_d = \dim_k
(k[x_{7, 2}, x_{7, 1}, x_{5, 1}, x_{3, 1}] / \initial(I))_d$ for 
all degrees $d \geq 0$. In particular,

\begin{align*}
	\dim_k \left(k[x_{7, 2}, x_{7, 1}, x_{5, 1}, x_{3, 1}] / I\right)_d &=
	\dim_k \left(k[x_{7, 2}, x_{7, 1}, x_{5, 1}, x_{3, 1}]	/ \initial(I)\right)_d \\
	&= \dim_k \left(k[x_{7, 2}, x_{7, 1}, x_{5, 1}, x_{3, 1}] / \langle
	x_{7, 2}^2, x_{5, 1}^2 \rangle \right)_d
\end{align*}
\noindent
for all $d \geq 0$. But the fact $\dim_k (R')_d = \dim_k \left(k[x_{7, 2},
x_{7, 1}, x_{5, 1}, x_{3, 1}] / \langle x_{7, 2}^2, x_{5, 1}^2
\rangle\right)_d$ can be checked by looking at the $k$-basis at each degree
$d$.

Therefore, the canonical ring $R'$ has presentation $R' =
k[x_{7, 2}, x_{7, 1}, x_{5, 1}, x_{3, 1}] / I$ with
initial ideal $\initial(I)$ generated by quadratics under grevlex
with $x_{3,1} \prec x_{5,1} \prec x_{7,1} \prec x_{7,2}$.
%These quadratics must also determine minimal generators for $I$ by
%Lemma ~\ref{lem:minimal_quadratic}.
Thus, $R'$ is generated up to degree $e = 7$ with relations up to
degree $2e = 14$, as desired.

@@

\begin{example}
\label{eg:base-1-33}
Let $(\sx',0 , \halfcan')$ be a log spin curve of genus $1$ with $\halfcan' = P - Q + \frac{1}{3}P_1 + \frac{1}{3}P_2$. In this example, we show
$$R_{\halfcan'} \cong k[u, x_3, y_3, y_4]/(x_3 y_3- \alpha uy_4, y_4^2 - \beta x_3^2 u - \gamma y_3^2u)$$
and that $(\sx', \halfcan', \{1,2\})$ is admissible.

@@

Then, there exist constants $\alpha, \beta, \gamma \in k$ so that 
$R_{\halfcan'} \cong k[u, x, y_3, y_4]/(xy_3- \alpha uy_4, y_4^2 - \beta x^2 u - \gamma y^2u).$ The proof of this is fairly algorithmic: We may first use Riemann--Roch to compute the dimensions of $(R_{\halfcan'})_n$ over $k$, then verify that these generators and relations produce the correct number of independent functions via an analysis of zero and pole order. The details are omitted as it is analogous to Example ~\ref{eg:base-0-377}.

@@

\begin{example}
\label{eg:exception-1-5}
Let $(\sx', 0, L')$ be a log spin curve of genus 1 with $L' = P -
 Q + \frac{2}{5} P_1.$ Let $x_2 \in (R_L)_2$ be any nonzero element.
We obtain $\di x_2|_{P_1} = 0,$ since $2P - 2Q \sim 0$ and by Riemann
--Roch, $\dim_k (P_L)_2 = 1.$ Let $y_3 \in (R_\halfcan)_3$ be any nonzero
element. We obtain $\di y_3|_{P_1} = - P_1,$ by Riemann Roch, since
if $y_3$ did not have a pole at $R$, we would obtain $y_3 \in H^0
(\sx,3P-3Q) \cong H^0(\sx, P - Q) \cong 0$ as $P \neq Q$. Finally,
let $y_5 \in (R_\halfcan)_5$ be an element with $y_5|_{P_1} = -P_1$. Then,
we claim there is some $\alpha \in k$ so that

$$R_\halfcan \cong k[x_2 , y_3, y_5]/(y_3^4 - \alpha x_2 y_5^2).$$

In order to show this is an isomorphism, one can use Riemann--Roch and pole order considerations at $P_1$ to check the above relation exists. One can then check that the generators and relation determine a ring $R$ so that for all $j \in \BZ,$ the $j$th graded component $(R)_j$ satisfies $\dim_k (R)_j = \dim_k R_\halfcan.$ Hence, $R \cong R_\halfcan$. The verification is analogous to Example ~\ref{eg:base-0-377} and is omitted.

@@

\item[(b)] Choose an ordering $\prec$ on $k[u, x_1, \ldots, x_m]$ such that
	\[
		\ord_u(f) < \ord_u(h) \implies f\prec h.
	\]
	Equip $k[y_1, \ldots, y_n]$ with graded $P$-lex, as defined in
	Definition ~\ref{defn:graded-p-lex} and $k[y_1, \ldots, y_n] \otimes k[u, x_1, \ldots, x_m]$ with block order.
	If $I'$ and $I$ are the ideals of relations of $k[u, x_1, \ldots, x_m]
	\to R_{\halfcan'}$ and $k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]
	\to R_{\halfcan}$ respectively, then

>>>>

\newcommand\nc{\newcommand}
\nc\on{\operatorname}
\nc\renc{\renewcommand}
\newcommand\ssec{\subsection}
\newcommand\sssec{\subsubsection}
\newcommand\BH{{\mathbb H}}
\newcommand\BN{{\mathbb N}}
\newcommand\BC{{\mathbb C}}
\newcommand\BF{{\mathbb F}}
\newcommand\BR{{\mathbb R}}
\newcommand\BQ{{\mathbb Q}}
\newcommand\BP{{\mathbb P}}
\newcommand\BZ{{\mathbb Z}}
\newcommand\BA{{\mathbb A}}
\newcommand\Bk{{\Bbbk}}

@@

Let $X$ be a smooth proper geometrically-integral algebraic curve of genus $g$ over a field $\Bk$.
It is well known that the canonical sheaf $\Omega _X,$ with
associated canonical divisor $K_X$, determines the {\bf canonical
map } $\pi: X \rightarrow \BP_\Bk^{g - 1}$. Then, the {\bf canonical
ring} is defined to be
\begin{align*}
	R(X, K_X) = \bigoplus_{d \geq 0} H^0(X, dK_X),
\end{align*}

@@

\begin{thm}
\label{thm:main}
Let $(\sx, \Delta, \halfcan)$ be a log spin curve over a perfect
field $\Bk$, so that $\sx$ has signature $\sigma = (g; e_1, \ldots,
e_r; \delta)$.

Then the log spin canonical ring is generated as a $\Bk$-algebra by 
elements of degree at most $e = \max(5, e_1, \ldots, e_r)$ with
relations generated in degrees at most $2e$,
so long as $\sigma$ does not lie in a finite list of exceptional
cases, as given in Table ~\ref{table:g-1-exceptional} for
signatures with $g = 1$ and Table ~\ref{table:g-0-exceptional} for
signatures with $g = 0$.
\end{thm}

\begin{rem}
\label{rem:main-weaker-assumptions}
In fact, the proof of Theorem ~\ref{thm:main} holds with $\Delta$ replaced by an arbitrary effective divisor of the coarse space.
 Furthermore, one may relax the assumption that $\Bk$ is perfect. Instead, one only need assume that the stacky curve
is separably rooted, as described further in Remark ~\ref{rem:stack-formalism}.
\end{rem}

@@

For the remainder of this paper, fix an algebraically closed field $\Bk$.
This is no restriction on generality, as generator and relation degrees are preserved under base change to the algebraic closure.

@@

\begin{defn}
\label{defn:stacky-curve}
A \textbf{stacky curve} $\sx$ over an algebraically closed field $\Bk$ is the datum of a smooth proper connected scheme $X$ of dimension $1$, together with a finite number of closed points of $X$, $P_1, \ldots, P_r$, called {\bf stacky points}, with {\bf stabilizer orders} $e_1, \ldots, e_r \in \BZ_{\geq 2}.$ The scheme $X$ associated to a stacky curve $\sx$ is called the {\bf coarse space} of $\sx$.

@@

The results of this paper can be easily couched in terms of the language of stacks. If one works over an arbitrary field $\Bk$ (which need not be algebraically closed) one can extend Theorem ~\ref{thm:main} to hold in the case that the stacky curve $\sx$ is tame and separably rooted, i.e.,~ the residue field of each
of the stacky points is separable.

@@

\begin{defn}
\label{defn:div-ex}
Let $\sx$ be a stacky curve over $\Bk$ with coarse space $X$ of genus $g$ and stacky points $P_1, \ldots,
P_r$ with stabilizer orders $e_1, \ldots, e_r \in \BZ_{\geq 2}$.
Then, we notate
\[
	\di \sx = \left(\bigoplus_{P\notin \{P_1, \ldots, P_r\}} \langle 
	P \rangle \right) \oplus \left(\bigoplus_{i = 1}^r \left \langle 
	\frac{1}{e_i}P_i \right \rangle \right) \subseteq \BQ \otimes \di X.
\]
\end{defn}

@@

\begin{defn}
\label{defn:h0-stacky}
Recall the standard notation 
$$\sco(D) = \{f \in \Bk(X)^\times : \di f + D \geq 0 \}.$$
Let $\sx$ be a stacky curve with coarse space $X$.
If $D \in \di \sx$ is a Weil divisor, then we define
\begin{align*}
	H^0(\sx, D) &:= H^0(X, \lfloor D \rfloor)\\
	H^0(\sx, \sco(D)) &:= H^0(\sx, D)\\
	h^0(\sx, \sco(D)) &:= \dim_\Bk H^0(\sx, \sco(D))
\end{align*}
If $R$ is a graded ring, then we let $(R)_d$ refer to the $d^\text{th}$
graded component of $R$.

@@

\begin{defn}
\label{defn:monomial-degree}
Let $\Bk[x_1$, $\ldots$, $x_n]$ be a graded polynomial
ring with $\deg x_i = d_i$ and let $\alpha := \prod_{i = 1}^{n} x_{i}^{f_i} \in
\Bk[x_{1}$, $\ldots$, $x_{n}]$ be a monomial. Then
we define the {\bf degree} of $\alpha$ to be
\[
	\deg \alpha := \sum_{i = 1}^{n} d_i f_i.
\]
\end{defn}

\begin{defn}
\label{defn:grevlex}
The {\bf graded reverse lexicographic order}, or {\bf grevlex} $\prec_{\text{grevlex}}$ is defined as follows.
If $\alpha := \prod_{i = 1}^{n} x_{i}
^{f_i}$ and $\beta := \prod_{i = 1}^{n} x_{i}^{f_i'}$ are
monomials in $\Bk[x_{1}$, $\ldots$, $x_{n}]$, then $\alpha
\succ_{\text{grevlex}} \beta$ if either

@@

\begin{defn}
\label{defn:block-order}
The \textbf{block term order} is defined as follows. Let $\Bk[y_{1}, \ldots, y_{m}]$ and $\Bk[x_{1}, \ldots,
x_{n}]$ be weighted polynomial rings with $\deg y_i = c_i, \: \deg x_i = d_i.$ Further assume we are given existing term
orders $\prec_y$ and $\prec_x$. Let $\alpha :=
\prod_{j = 1}^{m} y_{j}^{h_i}$ $\prod_{i = 1}^{n} x_{i}
^{f_i}$ and $\beta := \prod_{j = 1}^{m} y_{j}^{h_i'}
\prod_{i = 1}^{n} x_{i}^{f_i'}$ be monomials in $\Bk[y_{1},
\ldots, y_{m}] \otimes$ $\Bk[x_{1}, \ldots, x_{n}]$.
Let $\alpha_y := \prod_{j = 1}^{m} y_{j}^{h_i}$ be the
part of $\alpha$ in $\Bk[y_{1}, \ldots, y_{m}]$ and
likewise with $\alpha_x$, $\beta_y$, and $\beta_x$.

In the \textbf{(graded) block} (or \textbf{elimination}) term
ordering on $\Bk[y_{d_1', 1}, \ldots, y_{d_m', m}] \otimes \Bk[x_{d_1, 1},
\ldots, x_{d_n, n}]$, we define $\alpha \succ \beta$ if
  \begin{enumerate}
	\item[(i)] $\deg \alpha  > \deg \beta$ or
  \item[(ii)] $\deg \alpha = \deg \beta$ and $\alpha_y \succ_y \beta_y$ or
	\item[(iii)] $\deg \alpha = \deg \beta$ and $\alpha_y = \beta_y$ and $\alpha_x \succ_x \beta_x$.
  \end{enumerate}
\end{defn}

@@

\begin{defn}
\label{defn:initial-term}
Let $\prec$ be an ordering on $\Bk[x_{1}$, $\ldots$, $x_{n}]$, 
with $\deg x_i = d_i$,
and let $f \in \Bk[x_{1}, \ldots, x_{n}]$ be a homogeneous
polynomial. The \textbf{initial term} $\initial(f)$ of $f$
is the largest monomial in the support of $f$ with respect to
the ordering $\prec$. Furthermore, we set $\initial(0) := 0$.
\end{defn}

\begin{defn}
\label{defn:initial-ideal}
Let $I$ be a homogeneous ideal of $\Bk[x_{1}$, $\ldots$,
$x_{n}]$. Then the {\bf initial ideal} $\initial(I)$ of $I$ is
the ideal generated by the initial terms of homogeneous polynomials
in $I$:
\[
	\initial(I) := \langle \initial(f) \rangle_{f \in I}
\]
\end{defn}

@@

\begin{defn}
\label{defn:grobner-basis}
Let $I$ be a homogeneous ideal of $\Bk[x_{1}$, $\ldots$,
$x_{n}]$. A \textbf{Gr\"obner basis} for $I$, also known as a
\textbf{standard basis} for $I$, is a set of elements in $I$
such that their initial terms generate the initial
ideal of $I$.
\end{defn}

@@

Let $R'$ be the canonical ring associated to $\sx'$. Define $(\sx', \halfcan', J)$
to be {\bf admissible} if $R'$ admits a presentation
\begin{align*}
	R' \cong \left( \Bk[x_1, \ldots, x_m] \otimes \Bk[y_{i, e_i'}]_{i \in J} \right)/I'
\end{align*}

@@

\begin{example}
\label{eg:base-1-0}
Let $(\sx',0, \halfcan')$ be a log stacky curve of genus $g = 1,$ with $\halfcan' = 0$.
Counting dimensions, we see $h^0(\sx, dL')=1$ for all $d\in \mathbb{Z}_{\ge 0}$ so it is immediately clear that $R_{\halfcan'} \cong \Bk[x]$ with $x$ a generator in degree 1.
\end{example}

@@

Note that $R_{L_X} \cong \Bk[x_1, x_2]$ where $\deg x_1 = \deg x_2 = 1.
$ To see this, observe that we will need two generators in degree $1$
because $h^0(X, L_X) = 2$ by Riemann--Roch. Because $L_X$ is very 
ample, we have that $R_{L_X}$ is generated in degree 1. To
conclude, note that $R_{L_X}$ does not have any relations. If there
exists some relation, then $\dim \proj R_{L_X} < 1$. This would
contradict the fact that $L_X$ is very ample. Thus, $\proj R_{L_X}
\cong X$ which has dimension 1.

Next, we construct generators and relations for $R_L$ using those
of $R_{L_X}$. Note that we have a natural inclusion $\iota: R_{L_X}
\hookrightarrow R_L$ induced by the inclusions $H^0(\sx, dL_X)
\hookrightarrow H^0(\sx, dL)$ for each $d \geq 0$. By Riemann Roch,
we see there is some element $y_{1, 3} \in (R_L)_3$ with $\ord_{P_1}
(y_{1, 3}) = -1$, not in the image of the inclusion $\iota$. We
claim that there exist $a_1, a_2 \in \Bk$ and a degree $4$ polynomial
$f(x_1, x_2) \in \Bk[x_1, x_2]$ such that
\begin{align*}
	R_L \cong \Bk[x_1, x_2, y_{1, 3}] / (a_1 x_1 y_{1, 3} + a_2 x_2 y_{1, 3}
	+ f(x_1,x_2))
\end{align*}

@@

\noindent
are surjective for $d \geq 4$. A relation of the form $a_1 x_1 y_{1,
3} + a_2 x_2 y_{1, 3} + f(x_1, x_2) = 0$ must exist because $h^0(\sx,
4L) - h^0(\sx, 4L_X) = 1,$ but $x_1 y_{1, 3}$ and $x_2 y_{1, 3}$
define two linearly independent elements with nontrivial image in
the 1-dimensional vector space $H^0(\sx, 4L) / H^0(\sx, 4L_X).$ So,
we obtain a surjection
\begin{align}
\label{align:surjection}
	\Bk[x_1, x_2, y_{1, 3}] / (a_1 x_1 y_{1, 3} + a_2 x_2 y_{1, 3} + f(x_1, x_2)) 
	\rightarrow R_L.
\end{align}

To complete the example, it suffices to show there are no
additional relations. First, note that $a_1 x_1 y_{1, 3} + a_2 x_2
y_{1, 3} + f(x_1, x_2)$ is irreducible because there are no
relations among $x_1, x_2$ and $y_{1, 3}$ in lower degrees. Hence,
$\Bk[x_1, x_2, y_{1, 3}] / (a_1 x_1 y_{1, 3} + a_2 x_2 y_{1, 3} +
f(x_1, x_2))$ is integral and is 2-dimensional. Thus, the map
\eqref{align:surjection} defines a surjection from an integral
2-dimensional ring to a 2-dimensional ring. Therefore, it is an 
isomorphism.
\end{example}

@@

For $d = 0, 1, 2, \ldots$ we have
\[
	\dim_\Bk H^0 (\sx', d K_{\sx'}) = 1, 0, 0, 1, 0, 1, 1, 2, 1, 1, 2, 1, 3, 2, 3, \ldots
\]

@@

Let $I$ be the ideal generated by these relations in
$\Bk[x_{7, 2}, x_{7, 1}, x_{5, 1}, x_{3, 1}]$. Under grevlex with
$x_{3,1} \prec x_{5,1} \prec x_{7,1} \prec x_{7,2}$, the initial
ideal of $I$ is
\[
	\initial(I) = \langle x_{7, 2}^2, x_{5, 1}^2 \rangle
\]

@@

To demonstrate that these two relations generate all 
relations, we show that $\dim_\Bk (R')_d = \dim_\Bk
(\Bk[x_{7, 2}, x_{7, 1}, x_{5, 1}, x_{3, 1}] / \initial(I))_d$ for 
all degrees $d \geq 0$. In particular,

\begin{align*}
	\dim_\Bk \left(\Bk[x_{7, 2}, x_{7, 1}, x_{5, 1}, x_{3, 1}] / I\right)_d &=
	\dim_\Bk \left(\Bk[x_{7, 2}, x_{7, 1}, x_{5, 1}, x_{3, 1}]	/ \initial(I)\right)_d \\
	&= \dim_\Bk \left(\Bk[x_{7, 2}, x_{7, 1}, x_{5, 1}, x_{3, 1}] / \langle
	x_{7, 2}^2, x_{5, 1}^2 \rangle \right)_d
\end{align*}
\noindent
for all $d \geq 0$. But the fact $\dim_\Bk (R')_d = \dim_\Bk \left(\Bk[x_{7, 2},
x_{7, 1}, x_{5, 1}, x_{3, 1}] / \langle x_{7, 2}^2, x_{5, 1}^2
\rangle\right)_d$ can be checked by looking at the $\Bk$-basis at each degree
$d$.

Therefore, the canonical ring $R'$ has presentation $R' =
\Bk[x_{7, 2}, x_{7, 1}, x_{5, 1}, x_{3, 1}] / I$ with
initial ideal $\initial(I)$ generated by quadratics under grevlex
with $x_{3,1} \prec x_{5,1} \prec x_{7,1} \prec x_{7,2}$.
%These quadratics must also determine minimal generators for $I$ by
%Lemma ~\ref{lem:minimal_quadratic}.
Thus, $R'$ is generated up to degree $e = 7$ with relations up to
degree $2e = 14$, as desired.

@@

\begin{example}
\label{eg:base-1-33}
Let $(\sx',0 , \halfcan')$ be a log spin curve of genus $1$ with $\halfcan' = P - Q + \frac{1}{3}P_1 + \frac{1}{3}P_2$. In this example, we show
$$R_{\halfcan'} \cong \Bk[u, x_3, y_3, y_4]/(x_3 y_3- \alpha uy_4, y_4^2 - \beta x_3^2 u - \gamma y_3^2u)$$
and that $(\sx', \halfcan', \{1,2\})$ is admissible.

@@

Then, there exist constants $\alpha, \beta, \gamma \in \Bk$ so that 
$R_{\halfcan'} \cong \Bk[u, x, y_3, y_4]/(xy_3- \alpha uy_4, y_4^2 - \beta x^2 u - \gamma y^2u).$ The proof of this is fairly algorithmic: We may first use Riemann--Roch to compute the dimensions of $(R_{\halfcan'})_n$ over $\Bk$, then verify that these generators and relations produce the correct number of independent functions via an analysis of zero and pole order. The details are omitted as it is analogous to Example ~\ref{eg:base-0-377}.

@@

\begin{example}
\label{eg:exception-1-5}
Let $(\sx', 0, L')$ be a log spin curve of genus 1 with $L' = P -
 Q + \frac{2}{5} P_1.$ Let $x_2 \in (R_L)_2$ be any nonzero element.
We obtain $\di x_2|_{P_1} = 0,$ since $2P - 2Q \sim 0$ and by Riemann
--Roch, $\dim_\Bk (P_L)_2 = 1.$ Let $y_3 \in (R_\halfcan)_3$ be any nonzero
element. We obtain $\di y_3|_{P_1} = - P_1,$ by Riemann Roch, since
if $y_3$ did not have a pole at $R$, we would obtain $y_3 \in H^0
(\sx,3P-3Q) \cong H^0(\sx, P - Q) \cong 0$ as $P \neq Q$. Finally,
let $y_5 \in (R_\halfcan)_5$ be an element with $y_5|_{P_1} = -P_1$. Then,
we claim there is some $\alpha \in \Bk$ so that

$$R_\halfcan \cong \Bk[x_2 , y_3, y_5]/(y_3^4 - \alpha x_2 y_5^2).$$

In order to show this is an isomorphism, one can use Riemann--Roch and pole order considerations at $P_1$ to check the above relation exists. One can then check that the generators and relation determine a ring $R$ so that for all $j \in \BZ,$ the $j$th graded component $(R)_j$ satisfies $\dim_\Bk (R)_j = \dim_\Bk R_\halfcan.$ Hence, $R \cong R_\halfcan$. The verification is analogous to Example ~\ref{eg:base-0-377} and is omitted.

@@

\item[(b)] Choose an ordering $\prec$ on $\Bk[u, x_1, \ldots, x_m]$ such that
	\[
		\ord_u(f) < \ord_u(h) \implies f\prec h.
	\]
	Equip $\Bk[y_1, \ldots, y_n]$ with graded $P$-lex, as defined in
	Definition ~\ref{defn:graded-p-lex} and $\Bk[y_1, \ldots, y_n] \otimes \Bk[u, x_1, \ldots, x_m]$ with block order.
	If $I'$ and $I$ are the ideals of relations of $\Bk[u, x_1, \ldots, x_m]
	\to R_{\halfcan'}$ and $\Bk[u, x_1, \ldots, x_m, y_1, \ldots, y_n]
	\to R_{\halfcan}$ respectively, then
